Selective stock enhancement undermines community stability

### 2022-01-05 rmd/manuscript.Rmd
Update manuscript.Rmd

param_set <- df_theory %>% 
  filter(r_min == 0.5) %>% 
  distinct(n_species,
           k,
           r_type,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

@@

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species underpins emergent stability of ecological communities. In this context, stock enhancement represents excessive individuals that deviate from the internal regulation process of self-sustaining communities; therefore, it may disrupt the sensitive balance of species interactions with unforeseen consequences for community dynamics. Evidence for this hypothesis is lacking, however.

Here, we use a simple, general theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where \>10 millions of hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for this form of ecosystem management, reinforcing the recurrent calls for alternative management strategies.

## Theoretical prediction

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes the following values:

@@

In our simulation, we used XX parameter combinations that represent ecological scenarios with different strength of competition and XX. Under each scenario (i.e., parameter combination), we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with varying $R$. The last 1000 time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced strong overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.

>>>>

param_set <- df_theory %>% 
  filter(r_max == 2,
         alpha != 0.3) %>% 
  distinct(n_species,
           k,
           r_type,
           r1,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

@@

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild. Hence, this form of species management may have significant consequences for community dynamics. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time. Our theory illuminates that stock enhancement interferes the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche difference. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where \>10 millions of hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (interspecific competition drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$ whereas $\alpha_{ii}=1$), $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ takes the following values:

@@

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: mean strength of interspecific competition $\bar{\alpha}$, intrinsic growth of the enhanced species $r_1$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario (i.e., parameter combination), we ran 1400 time steps of 1000 independent communities (i.e., simulation replicates) with varying $R$. The last 1000 time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, the reduced mean of unenhanced species is caused by interspecific competition between enhanced and unenhanced species. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.
